---
title: "R Notebook"
output: html_notebook
---

Import all libraries 

```{r}
library(dplyr)
library (ggplot2)
```

Configure settings
```{r}
# disable scientific notation
options(scipen=999)

# disable warnings
options(warn = - 1) 

```



```{r}
#Establish connection to the database
library(RMySQL) 
con = dbConnect(MySQL(), dbname = "BnS",
                host = "mysql-1.cda.hhs.se", port = 3306,
                user = "bns", password = "bns@sse")


# Query the receipts data set

receipt_allocation <- dbGetQuery(con, 
 "SELECT 
    receipt_id,
    MAX(is_online) AS is_online
  FROM Transactions 
    LEFT JOIN Unseen 
    USING (receipt_id)
  WHERE id23500 = 1 ##change to your studentid
    OR is_online IS NOT NULL
  GROUP BY receipt_id")

all_transactions <- dbGetQuery(con, 
  "SELECT *
  FROM Transactions"                        
                           )

products <- dbGetQuery(con, 
  "SELECT *
  FROM Products"                        
                           )

customers <- dbGetQuery(con, 
  "SELECT *
  FROM Customers"                        
                           )


# close the connection to db
lapply( dbListConnections( dbDriver( drv = "MySQL")), dbDisconnect)
```

Here is a code snippet confirming that all transactions in a receipt have the
same is_online status

```{r}
library(dplyr)

transactions %>%
  group_by(receipt_id) %>%
  mutate(vec_len = length(unique(is_online))) %>%
  mutate(transactions = length(receipt_id)) %>%
  mutate(total = sum(amount)) %>%
  filter(vec_len != 1)
  

```


A final overview check for missing values. Which varaibles would I have to include
in order to do imputation

```{r}

summary(products)
full_db <- left_join(left_join(transactions, products, by = "item"), customers, by = "customer_id")

full_db <- full_db %>%
  mutate(
    gender = as.factor(gender),
    enrollment = as.factor(enrollment),
    city = as.factor(city)
  )

summary(full_db %>% filter(is_online == 1))

```





Prepare the transactions table for id23500

```{r}

transactions <- all_transactions %>%
  filter(receipt_id %in% receipt_allocation$receipt_id) %>%
  mutate(
    purchase_date = as.Date(purchase_date,"%Y-%m-%d"),
    purchase_month = as.factor(format(purchase_date, "%Y-%m")),
    is_online = as.factor(is_online)
    )

transactions <- left_join(transactions, products, by = "item")


# before we join the customer data, lets check the completeness of customer_id 
# in the transaction table

# Check if there are any invalid customer_ids
transactions %>% 
  filter(!(customer_id %in% customers$customer_id ))

# Correct by converting the invalid customer_id to numeric
transactions <- transactions %>%                               # Replacing values
  mutate(customer_id = replace(customer_id, customer_id == "01-nov-00", 
                               as.numeric(as.Date("01-nov-00", "%d-%b-%y"))))

# confirm that the customer_id has been correctly updated
transactions %>%
  filter(receipt_id == "382324908696576")


transactions <- left_join(transactions, customers, by = "customer_id")

# look at a summary of the data
summary(transactions)

# Investigate how many missing values we have
sapply(transactions, function(x) sum(is.na(x)))
# we have 1 customer_id in the transactions table that is not found in the customers table
# we have 2 rows with item(s) in the transactions table that are not found in the products table



# Investigate the invalid item
transactions %>% 
  filter(!(item %in% products$item ))  # not problem of item not in products table

na_items <- unique((transactions %>% filter(is.na(dept)))[,'item'])
products %>% filter(item %in% na_items )  # items have NA in the products table




# Impute the values of the missing item AH0000

length(unique(transactions$receipt_id))
length(unique(transactions$item))
length(unique(transactions$subclass))

length(unique(transactions$class))


unique(customers$enrollment)


```

Now we will cluster the transactions on receipt level

```{r}


modes <- function(x) {
   ux <- unique(x)
   tab <- tabulate(match(x, ux))
   ux[tab == max(tab)]
}

head(transactions)



receipts <- transactions %>% 
  group_by(receipt_id) %>%
  summarize(
    customer_id = customer_id[1],
    purchase_date = purchase_date[1],
    purchase_month = purchase_month[1],
    tot_quantity = sum(quantity), 
    tot_amount = sum(amount),
    is_online = is_online[1], 
    mode_dept = modes(dept)[1], 
    age = age[1], 
    gender = gender[1],
    city = city[1], 
    enrollment = enrollment[1]
    )

library(tidyverse)
# make some format changes
receipts <- receipts %>%
  mutate(
    is_online = fct_recode(as.factor(is_online), offline = "0", online = "1"), 
    
    )



receipts

glimpse(receipts)

```

Now we will explore AGE on the receipts aggregation level


```{r}
## LOOKING AT AGE

# gender distribution within online offline
receipts %>%
  filter(!is.na(is_online)) %>%
  group_by(is_online, gender) %>%
  summarize(
    count = length(is_online)
  ) %>%
  mutate(pct = 100 * count / sum(count))
 

receipts %>%
  filter(!is.na(is_online)) %>%
  group_by(gender, is_online) %>%
  summarize(
    count = length(is_online)
  ) %>%
  mutate(pct = 100 * count / sum(count))


receipts %>%
  filter(!is.na(is_online)) %>%
  group_by(is_online) %>%
  summarize(
    count = length(is_online)
  ) %>%
  mutate(pct = 100 * count / sum(count))
 

# Gender distribution of all of the data where we know online/offline
receipts %>%
  filter(!is.na(is_online)) %>%
  group_by(gender) %>%
  summarize(
    count = length(is_online)
  ) %>%
  mutate(pct = 100 * count / sum(count))
 
# Gender distribution of the entire dataset
receipts %>%
  group_by(gender) %>%
  summarize(
    count = length(is_online)
  ) %>%
  mutate(pct = 100 * count / sum(count))

```


Now we will explore AGE on the receipts aggregation level



```{r}

# group ages into bins
generations <- seq(0, 100, 10)
for (g in generations) {
  receipts <- receipts %>%
  mutate (
    generation = ifelse(age >= g & age < (g+10), g, generation)
  )
}

receipts

# Age distribution within online offline
receipts %>%
  filter(!is.na(is_online)) %>%
  group_by(generation, is_online) %>%
  summarize(
    count = length(is_online)
  ) %>%
  mutate(pct = 100 * count / sum(count))


# Age distribution all purchases controlling filtering out is_online NAs
# Q: What is the age distribution of all our customers
receipts %>%
  filter(!is.na(is_online)) %>%
  group_by(generation) %>%
  summarize(
    count = length(is_online)
  ) %>%
  mutate(pct = 100 * count / sum(count))


# Q: What is the age adn gender distribution of all our customers
receipts %>%
  filter(!is.na(is_online)) %>%
  group_by(generation, gender) %>%
  summarize(
    count = length(is_online)
  ) %>%
  mutate(pct_in_bin = 100 * count / sum(count)) %>%
  ungroup() %>%
  mutate(pct = 100 * count / sum(count))

# Q What is the most common age and sex of customer?
receipts %>%
  filter(!is.na(is_online)) %>%
  group_by(generation, gender) %>%
  summarize(
    count = length(is_online)
  ) %>%
  mutate(pct_in_bin = 100 * count / sum(count)) %>%
  ungroup() %>%
  mutate(pct = 100 * count / sum(count)) %>%
  arrange(desc(pct))
  


# Q What if we add online/offline?
receipts %>%
  filter(!is.na(is_online)) %>%
  group_by(generation, is_online) %>%
  summarize(
    count = length(is_online)
  ) %>%
  mutate(pct_in_bin = 100 * count / sum(count)) %>%
  ungroup() %>%
  mutate(pct = 100 * count / sum(count)) %>%
  arrange(desc(is_online), desc(pct))
  


# Q Where do the online revenues come from?
receipts %>%
  filter(!is.na(is_online)) %>%
  group_by(generation, gender, is_online) %>%
  summarize(
    count = length(is_online)
  ) %>%
  mutate(pct_in_bin = 100 * count / sum(count)) %>%
  ungroup() %>%
  mutate(pct = 100 * count / sum(count)) %>%
  arrange(desc(is_online), desc(pct))
  

```



Lets see if there is a diff in tot_quantity or tot_amount when online/offline

```{r}

receipts
receipts %>%
  filter(!is.na(is_online)) %>%
  group_by(is_online) %>%
  summarize(
    avg_quantity = mean(tot_quantity),
    adj_sd_quantity = sd(tot_quantity) / avg_quantity,
    avg_amount = mean(tot_amount),
    adj_sd_amount = sd(tot_amount) / avg_amount
  ) 



```

Now we will explore DEPT on the receipts aggregation level


```{r}

length(unique(transactions$dept))
length(unique(transactions$class))
length(unique(transactions$subclass))



transactions %>%
  group_by(city) %>%
  summarize(count = length(receipt_id))


receipts %>%
  filter(!is.na(is_online)) %>%
  group_by(mode_dept) %>%
  mutate(count_mode_dept = length(mode_dept)) %>%
  ungroup() %>%
  group_by(mode_dept, is_online) %>%
  summarize(
    count = length(mode_dept),
    share = 100 * count / count_mode_dept[1]
  ) %>%
  #arrange(mode_dept, is_online)
  arrange(desc(is_online), desc(share))
  


# What is the baseline  = 7.3%
receipts %>%
  filter(!is.na(is_online)) %>%
  mutate(count_mode_dept = length(mode_dept)) %>%
  group_by(is_online) %>%
  summarize(
    count = length(mode_dept),
    share = 100 * count / count_mode_dept[1]
  ) %>%
  #arrange(mode_dept, is_online)
  arrange(desc(is_online), desc(share))



```


Now we will explore ENROLLMENT on the receipts aggregation level


```{r}


# Enrollment Baseline
# Q: how are all receipts distributed across enrollment forms?
receipts %>%
  filter(!is.na(is_online)) %>%
  mutate(tot_count = length(enrollment)) %>%
  group_by(enrollment) %>%
  summarize(
    count = length(enrollment),
    share = 100 * count / tot_count[1]
  ) %>%
  arrange(desc(share))


# Online Offline Baseline
# How are all receipts distributed across being online or offline?
receipts %>%
  filter(!is.na(is_online)) %>%
  mutate(count_mode_dept = length(mode_dept)) %>%
  group_by(is_online) %>%
  summarize(
    count = length(mode_dept),
    share = 100 * count / count_mode_dept[1]
  ) %>%
  #arrange(mode_dept, is_online)
  arrange(desc(is_online), desc(share))



# Online Offline across Enrollment
# Q: which enrollment forms have higher online activity than baseline above?
receipts %>%
  filter(!is.na(is_online)) %>%
  group_by(enrollment) %>%
  mutate(tot_count = length(is_online)) %>%
  ungroup() %>%
  group_by(is_online, enrollment) %>%
  summarize(
    count = length(is_online),
    share = 100 * count / tot_count[1]
  ) %>%
  arrange(desc(is_online), enrollment)
# Results:
# 7.3% of receipts online is the benchmark
# 59% of ECOM enrollment are online


# Enrollment across Online Offline
receipts %>%
  filter(!is.na(is_online)) %>%
  group_by(is_online) %>%
  mutate(tot_count = length(is_online)) %>%
  ungroup() %>%
  group_by(is_online, enrollment) %>%
  summarize(
    count = length(is_online),
    share = 100 * count / tot_count[1]
  ) %>%
  arrange(desc(is_online), desc(share))
  




```


Look at all of the opting out variables 

```{r}

customers2 <- customers %>%
  mutate(
    no_contact = as.factor(no_contact), 
    no_email = as.factor(no_email), 
    no_fax = as.factor(no_fax), 
    no_mail = as.factor(no_mail), 
    no_sms = as.factor(no_sms), 
    no_panel = as.factor(no_panel)
    ) %>%
  select(customer_id, no_contact, no_email, no_fax, no_mail, no_sms, no_panel)

receipts2 <- left_join(receipts, customers2, by = "customer_id")
head(receipts2)



# check the impact of no_xxx on online/offline

# Enrollment across Online Offline
receipts2 %>%
  filter(!is.na(is_online)) %>%
  group_by(is_online) %>%
  mutate(tot_count = length(is_online)) %>%
  ungroup() %>%
  group_by(is_online, no_panel) %>%
  summarize(
    count = length(is_online),
    share = 100 * count / tot_count[1]
  ) %>%
  arrange(desc(is_online), desc(share))
  

summary(customers2)

```





Are there any free transactions 

```{r}

# are there any free transactions?
# which free items are most common>
receipts %>% 
  group_by(item) %>%
  mutate(
    total_transactions = length(item)
  ) %>%
  ungroup() %>%
  filter(tot_amount == 0) %>%
  group_by(item) %>%
  summarize(
    free_transactions = length(item),
    total_transactions = max(total_transactions)
  ) %>%
  arrange(desc(free_transactions))

receipts %>%
  filter(tot_amount == 0) %>%
  arrange(desc(tot_quantity))

```



Confirmed that receipts have one instance of every item, with the exception of
one receipt. 

```{r}

# summary statistics of item transactions
transactions %>%
  group_by(item) %>%
  summarize(
    tot_quantity=sum(quantity),
    tot_transactions = length(receipt_id),
    tot_receipts = length(unique(receipt_id)),
    diff = abs(tot_transactions - tot_receipts)
    ) %>%
  arrange(desc(diff), desc(tot_quantity))

transactions

# lets see if we can find the receipt with two transactions of item 36519
transactions %>%
  filter(item == "36519") %>%
  group_by(receipt_id) %>%
  summarize(occurance = length(receipt_id)) %>%
  arrange(desc(occurance))

# we see that receipt has two occurrences of 568929403185978

transactions %>%
  filter(receipt_id == "568929403185978")


```


Look at the distribution of quantites and amounts(revenues)?
Turns out that this got too specific. We should look on class level

```{r}

# summary statistics of item transactions
items <- transactions %>%
  group_by(item) %>%
  summarize(
    tot_quantity=sum(quantity),
    tot_transactions = length(receipt_id),
    tot_amount = sum(amount),
    avg_price = mean(amount/quantity)
    ) %>%
  arrange(desc(tot_amount))

items

# check for negative item codes
#sort(unique(transactions[,'item']))


  
```



How do our transactions very on date?
Make a histogram showing the revenue for each month, 
split by color online or offline


```{r}

monthly_data <- transactions %>%
  group_by(purchase_month) %>%
  # UNTIL WE HAVE COMPLETED IMPUTATIONS, WE DROP MISSING IS_ONLINE
  filter(is.na(is_online) == FALSE) %>%
  mutate(
    online_amount = ifelse(is_online == 1, amount, 0)
  ) %>%
  summarize(
    tot_quantity = sum(quantity),
    tot_online = sum(is_online),
    tot_amount = sum(amount), 
    tot_amount_online = sum(online_amount)
  )

monthly_data


library(ggplot2)

# Look at the distribution of REVENUE
ggplot(transactions, aes(x = purchase_month, weight = amount)) +
  geom_bar(aes(fill = is_online)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(x = "Month", y = "Revenue (Million)") +
  scale_y_continuous(labels = unit_format(unit = "", scale = 1e-6)) + 
  scale_fill_discrete(name = "", labels = c("In Store", "Online", "NA")) +
  ggtitle("BnS Point of Sale Data - Revenue")

# Look at the distribution of QUANTITY
ggplot(transactions, aes(x = purchase_month, weight = quantity)) +
  geom_bar(aes(fill = is_online)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(x = "Month", y = "Quantity") +
  scale_fill_discrete(name = "", labels = c("In Store", "Online", "NA")) +
  ggtitle("BnS Point of Sale Data - Quantity")


```



Now we will investigate the relationship between online and 

```{r}










```







## Logistic Regression

The first model that I will evaluate is a logistic regression with 10-fold cross-validation.

```{r eval=FALSE, include=FALSE}
# define training control
train_control <- trainControl(method = "cv", number = 10)

# train the model on training set
online_model_LR <- train(fmla,
               data = df,
               trControl = train_control,
               method = "glm",
               family=binomial())

online_model_LR$results

# print cv scores
summary(model)


```





Excess

```{r eval=FALSE, include=FALSE}



### PLOT OF TREEMAP
# plot a pie chart of revenues
#library(plotly)
#library(treemapify)
#library(ggplot2)
# 
#p <- ggplot(items %>% slice_head(n = 5), aes(area = tot_amount, fill = item)) +
#  geom_treemap()
#
#p


length(customers$customer_id) #22,289
length(unique(customers$customer_id))
length(customers$host_p_id) # 22,289
length(unique(customers$host_p_id)) #390

length(unique(transactions$customer_id)) #22,138

# chekc the unique transactions against customer_id
combined <- unique(c(customers$customer_id, transactions$customer_id))
length(combined) #22,290
setdiff(combined, customers$customer_id) 

# check the unique transactions against host_p_id
combined <- unique(c(customers$host_p_id, transactions$customer_id))
length(combined) #22,141
setdiff(combined, customers$customer_id) 

```

